Magnetic Inhibition of Convection and the Fundamental Properties of Low--Mass Stars

### 2016-02-22 paper/ms.tex
revised and expanded conclusions + acknowledgements

\section{Summary \& Conclusions}
\label{sec:tellit}
Stellar evolution models and literature data are used to derive ages for stars with spectral types A through M in Upper Scorpius. I find a revised median HR diagram age of {\bf 9 -- 10} Myr for A-, F-, and G-type stars, with the exception of a population of F-type stars that appear 5 Myr older. These results agree with the revised $11\pm3$ Myr of \citet{Pecaut2012}. At the same time, I confirm the HR diagram age of 4 -- 5 Myr for K- and M-type stars using standard stellar evolution isochrones \citep{Preibisch2012, Slesnick2008, Herczeg2015}. However, isochrone ages of K- and M-type stars in eclipsing multiple-star systems observed by \emph{Kepler}/K2 \citep{Kraus2015, Alonso2015} disagree with the HR diagram analysis and suggest ages between 6 and 8 Myr for the same stars.

Application of stellar evolution models accounting for magnetic inhibition of convection \citep{FC12b, FC13} yields an age of {\bf 9 -- 10} Myr for K- and M-type stars in Upper Scorpius. Magnetic models provide a consistent explanation for the age discrepancy between high- and low-mass stars, as well as differences observed between HR diagram and mass-radius relationship ages. Magnetic models also reproduce the slope of the low-mass star mass-radius relationship. Critically, average surface magnetic field strengths are determined \emph{a priori} via equipartition arguments, highlighting the potential predictive power of magnetic stellar models of young stars. Furthermore, magnetic models around 10 Myr exhibit a lithium depletion pattern that is in agreement with inferences from observed lithium equivalent widths \citep{Rizzuto2015}.

Results from this work and \citet{Pecaut2012} strongly support adoption of an approximately 10 Myr median age for the Upper Scorpius subgroup of the Scorpius-Centaurus OB association. There is no population of stars in the Upper Scorpius that unambiguously support an age of 5 Myr. Still, additional eclipsing binary systems are needed to further populate the mass-radius relationship and solidify Upper Scorpius as a 10 Myr old stellar association.

Further work is needed to explore the impact of magnetic inhibition of convection on the (sub)stellar initial mass function for Upper Scorpius and other young stellar populations \citep[e.g.,][]{Ardila2000, Preibisch2002}. More work is also needed to explore whether magnetic inhibition of convection serves as a robust solution for noted age gradients as a function of \teff\ \citep{Naylor2009, Herczeg2015}. However, given the ease with which magnetic models appear to reconcile discrepancies in Upper Scorpius, the wider application of magnetic stellar models as a component of isochrone analyses is recommended. A small grid of magnetic stellar isochrones is under construction as is a web interface to request custom magnetic models.

\begin{acknowledgements}
Thank you to Alexis Lavail, Eric Mamajek, James Silvester, and Eric Stempels for reading and commenting on an early version of this manuscript. The magnetic Dartmouth stellar evolution code was originally developed with support of National Science Foundation (NSF) grant AST-0908345. This work made use of NASA's Astrophysics Data System (ADS) and the VizieR catalogue access tool by CDS in Strasbourg, France. Figures in this manuscript were produced with Gnuplot 5 \citep{Gnuplot5.0}.
\end{acknowledgements}

>>>>

\section{Summary \& Conclusions}
\label{sec:tellit}
Stellar evolution models and literature data are used to {\bf estimate} median ages for stars with spectral types A through M in Upper Scorpius. I find a median HR diagram age of {\bf 9 -- 10} Myr for A-, F-, and G-type stars, with the exception of a population of F-type stars that appear 5 Myr older. These results agree with the revised $11\pm3$ Myr {\bf proposed by} \citet{Pecaut2012}. At the same time, I confirm the HR diagram age of 4 -- 5 Myr for K- and M-type stars using standard stellar evolution isochrones \citep{Preibisch2012, Slesnick2008, Herczeg2015}. {\bf Notably}, isochronal ages of K- and M-type stars in eclipsing multiple-star systems observed by \emph{Kepler}/K2 \citep{Kraus2015, Alonso2015, David2015b} disagree with the HR diagram analysis and suggest ages between 6 and 8 Myr for the same stars.

Using stellar evolution models {\bf that account} for magnetic inhibition of convection \citep{FC12b, FC13}, {\bf I find} an age of {\bf 9 -- 10} Myr for K- and M-type stars in Upper Scorpius. Magnetic inhibition of convection consistently explains:
\begin{enumerate} 
	\item the age discrepancy between high- and low-mass stars in the HR diagram,
	\item the observed slope of the low-mass stellar mass-radius relationship,
	\item stellar age differences observed between HR diagram and mass-radius relationship determinations,
	\item and the lithium depletion pattern inferred from observed lithium equivalent widths \citep{Rizzuto2015}. 
\end{enumerate}
Critically, average surface magnetic field strengths are determined \emph{a priori} via equipartition arguments, highlighting the potential predictive power of magnetic stellar models for young stars.

Results from this work and \citet{Pecaut2012} strongly support the adoption of an approximately 10 Myr median age for the Upper Scorpius subgroup of the Scorpius-Centaurus OB association. There is no population of stars in the Upper Scorpius that unambiguously supports an age of 5 Myr. {\bf Nevertheless, additional eclipsing binary systems are needed to better populate the mass-radius relationship to solidify Upper Scorpius as a 10 Myr old stellar association and to verify predictions from magnetic stellar evolution models across the full mass spectrum.}

{\bf Further work is needed to explore whether magnetic inhibition of convection serves as a robust solution for noted age gradients as a function of stellar spectral type \citep[e.g.,][]{Naylor2009, Herczeg2015}. Previous work determining ages for the $\beta$-Pictoris moving group from magnetic stellar models suggests the solution is robust \citep{MM10, Malo2014, Binks2016}, but additional confirmation using stellar populations with a wider variety of ages is required. These results also highlight the need to explore how the inclusion of magnetic inhibition of convection affects the (sub)stellar initial mass function young stellar populations.}

In the meantime, the use of magnetic stellar models as a component for isochrone analyses for young low-mass stars is recommended. Consistency across the HR and mass-radius diagrams exhibited by magnetic stellar evolution isochrones suggest that magnetic isochrones will provide a more accurate characterization of transiting exoplanet host stars in young stellar populations, especially those observed by \emph{Kepler}/K2 in Upper Scorpius. By adopting an incorrect age, a reasonable estimate of stellar radii can be found, but stellar masses will be systematically underestimated by roughly 30\%. Similar arguments apply to the characterization of young low-mass star hosting directly imaged giant planets or brown dwarfs. Ages inferred from standard model isochrones will be systematically too young by a factor of two resulting in a systematic underestimation of giant planet and brown dwarfs masses by roughly 35\%. To facilitate this work, all models used in this study are available online\footnote{\url{https://github.com/gfeiden/MagneticUpperSco/}}. In addition, a small grid of solar-metallicity magnetic stellar mass tracks and isochrones is under construction, as is a web-interface to compute custom magnetic models.

\begin{acknowledgements}
G.A.F. thanks Alexis Lavail, Eric Mamajek, James Silvester, and Eric Stempels for reading and commenting on an early version of this manuscript as well as Bengt Gustafsson and Thomas Nordlander for interesting discussions that helped improved the manuscript. The magnetic Dartmouth stellar evolution code was originally developed with support of National Science Foundation (NSF) grant AST-0908345. This work made use of NASA's Astrophysics Data System (ADS), the VizieR catalogue access tool by CDS in Strasbourg, France \citep{Ochsenbein2000}, and IPython with Jupyter notebooks \citep{Perez2007}. Figures in this manuscript were produced with Gnuplot~5 \citep{Gnuplot5.0}.
\end{acknowledgements}
